---
title: Motor Trend Analysis | Relationship between a set of variables and miles per
  gallon (MPG) (outcome).
author: "Roxana Trejos Ramírez"
date: "11/27/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
Motor Trend is looking at a data set of a collection of cars, they want to explore the relationship between a set of variables and miles per gallon (MPG) (outcome) by answering:
A.  Is an automatic or manual transmission better for MPG?
B.  Quantify the MPG difference between automatic and manual transmissions.
In this research, we are clarifying those questions.

# 1.  Processing Data and Exploring Data
```{r include=FALSE}
    library(datasets)
    library(ggplot2)
    data(mtcars)
```
## 1.1  Headers of MTCARS data
```{r}
    # [, 1]	mpg	Miles/(US) gallon
    # [, 2]	cyl	Number of cylinders
    # [, 3]	disp	Displacement (cu.in.)
    # [, 4]	hp	Gross horsepower
    # [, 5]	drat	Rear axle ratio
    # [, 6]	wt	Weight (1000 lbs)
    # [, 7]	qsec	1/4 mile time
    # [, 8]	vs	Engine (0 = V-shaped, 1 = straight)
    # [, 9]	am	Transmission (0 = automatic, 1 = manual)
    # [,10]	gear	Number of forward gears
```
## 1.2  Internal Structure of MTCARS
```{r include=FALSE}
    str(mtcars)
```
#  2.  Questions
## IS AN AUTOMATIC OR MANUAL TRANSMISSION BETTER FOR  MPG?
###  Analyzing the transmision performance of MTCARS 
By using SUMMARY function we are analyzing the statistical results of MTCARS.
```{r}
# A.  Checking data from summary.
#       0 = automatic
summary(mtcars[mtcars$am==0,])
#       1 = manual
summary(mtcars[mtcars$am==1,])
```
###  Testing the hypothesis: *Manual transmission is better than automatic transmission for MPG*
Analyzing the data of MTCARS and trying to proof the hypothesis.
```{r}
    t.test(mtcars$mpg~mtcars$am)
```
**CONCLUSIONS:**

1.  The mean of mpg is greater for manual (at 24.4) than automatic (at 17.1).

2.  Null hypothesis rejected.
Based on the results, 
**p-value** = 0.001374<0.05, 
therefore, 
a.  We reject the null hypothesis that there is no difference in MPG, 
b.  Conclude that:
-  Manual transmission is better than automatic transmission for MPG, 
-  Assumption that all other conditions remain unchanged.
By analyzing the mean of MPG, the general conclusion is that manual vehicles have better        
MPG performance than Automatic vehicles. 
But by how much? In the following analysis we will investigate the reason why manual is         better than automatic by quantifying the difference.

## QUANTIFY THE MPG DIFFERENCE BETWEEN AUTOMATIC AND MANUAL TRANSMISSIONS
###  Analyzing the transmision performance of MTCARS graphically.
By visualizing the data of the MTCARS.
```{r include=FALSE}
    boxplot(mpg ~ am, 
            data = mtcars, 
            xlab = "Transmission (0 = Manual and 1 = Automatic)", 
            ylab = "Miles per gallon", 
            main="Miles per gallon by Transmission Type")
```    
Manual has a higher mean for mpg than automatic.
### Analyzing the linear model
```{r include=FALSE}
    stepModel = step(lm(data = mtcars, mpg ~ .),trace=0,steps=10000)
    summary(stepModel)
```
Our model has included three variables:

- **wt** - Weight (1000 lbs)
- **qsec** - qsec	1/4 mile time
- **am** - Transmission (0 = automatic, 1 = manual)

This model explains 85% of the variance. It may be concluded that on average, manual transmissions have 1.478 more mpg than automatic.
To optimize the model, let's examine the model with mpg ~ wt + qsec correlation with am.
```{r include=FALSE}
    model <- lm(mpg~ factor(am):wt + factor(am):qsec,data=mtcars)
    summary(model)
```
This model has a 89.5% total variance with an adjusted variance of 88%. 
By adding the coefficients, we have the following conclusions:

- When the weight increased by 1000 lbs, the mpg decreased by -3.176 for automatic transmission cars, and -6.09 for manual transmission cars
-  By increasing car weight,  the proper transmission to choose is:  Manual.
-  When the acceleration speed dropped and 1/4 mile time increased by 1 sec:
   + The mpg factor increased by 0.834 miles for automatic transmission cars.
   + And,  1.446 miles for manual transmission cars.
-  Lowering acceleration speed, but same weight, manual transmission cars are better for mpg.

## MAIN CONCLUSION
The mpg is determined by the interaction among: weight, acceleration and transmission. Given the above analysis, the original question (automatic transmission vs manual transmission) is not  answered, and we should consider the context of weight and acceleration speed (clarified in question 2).

#  Appendix

## Apendix 1:  Residual check and diagnostics
```{r}
    par(mfrow=c(2,2))    
    plot(model)
```

## Apendix 2:  Regression Model Results
```{r}
    full.model <- lm(mpg ~ ., data = mtcars)
    best.model <- step(full.model, direction = "backward")
```